\documentclass[a4paper]{article}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}

\title{Advanced Differentials Final Paper}

\author{Arleigh Dickerson}

\date{\today}

\begin{document}
\maketitle

\section{Introduction}
Hello, friends. Today we will be learning about trivial and nontrivial solutions to Floquet systems and how the existence of a nontrivial solution relates to the system's Floquet multipliers. Let's begin.
\section{Theorem 2.72}

In this theorem, we see some of the qualitative properties of the trivial solution. Specifially, this theorem examines how the value of Floquet multipliers influences the trivial solution's stability or lack thereof. To accomplish this end, we will and link our claim to the Stability Theorem (2.49).

\subsection{Theorem}

Let $\mu_1$, $\mu_2$,\ldots, $\mu_n$ be the Floquet multipliers of the Floquet system $x' = A(t)x$. Remember from previously that for a $\mu$ to be a Floquet multiplier of a system, it must be an eigenvalue of $C := \Phi^{-1}(0)\Phi(\omega)$. Our claim is that the trivial solution is \begin{enumerate}
    \item globally asymptotically stable on $[0,\infty)$
    \item stable on $[0,\infty)$ provided $|\mu_i| \leq 1$, $1 \leq i \leq \infty$ and whenever $|\mu_i| = 1$, $\mu_i$ is a simple eigenvalue
    \item unstable on $[0,\infty)$ provided there is an $i_0$, $1 \leq i_0 \leq n$ such that $|\mu_{i_0}| > 1$.
\end{enumerate}

\subsection{Proof}
The following proof is for the two-dimensional case. In proving Floquet's theorem, recall that we chose an arbitrary $B$ such that $e^{B \omega} = C$. 
Also, notice that the Jordan canonical form theorem gives us matrices $M$ and $J$ such that $B = MJM^{-1}$, where either $J = \begin{bmatrix}
    \rho_1 & 0 \\
    0 & \rho_2
\end{bmatrix}$ or $J = \begin{bmatrix}
    \rho_1 & 1 \\
    0 & \rho_1
\end{bmatrix}$ where $\rho_1$, $\rho_2$ are the eigenvalues of $B$. Finally, remember that if a matrix $Y$ is invertible, then $e^{Y X Y^{-1}} = Y e^{X} Y^{-1}.$
With these properties in mind, it follows that
\begin{equation*}
\begin{split}
    C &= e^{B\omega}\\
      &= e^{MJM^{-1}}\\
      &= M e^{J\omega} M^{-1}\\
      &= M K M^{-1}\\
\end{split}
\end{equation*}
where either $K = \begin{bmatrix}
    e^{\rho_{1}\omega} & 0 \\
    0 & e^{\rho_{2}\omega}\\
\end{bmatrix}$, or $K = \begin{bmatrix}
    e^{\rho_{1}\omega} & \omega e^{\rho_{1}\omega}\\
    0 & e^{\rho_{1}\omega}\\
\end{bmatrix}$.

Because the eigenvalues of $K$ are the same as the eigenvalues of $C$, we can see that the Floquet multipliers are $\mu_i = e^{\rho_{i}\omega}$ with $i = 1,2$ given that it is possible that $\rho_1 = \rho_2$. With the fact that $|\mu_i| = e^{Re(\rho_i)\omega}$ in mind, we show that 
\begin{equation*}
\begin{split}
    |\mu_i| < 1 &\text{ iff } Re(\rho_i) < 0 \\
    |\mu_i| = 1 &\text{ iff } Re(\rho_i) = 0 \\
    |\mu_i| > 1 &\text{ iff } Re(\rho_i) > 0\text{,} \\
\end{split}
\end{equation*} where the $Re$ function discards the complex portion of a value and returns only the real part.

As explained by Kelly, the equation from Theorem 2.71 
\begin{equation*}
    x(t) = P(t)y(t)
\end{equation*}
establishes a bijection between solutions of the Floquet system $x' = A(t)x$ and $y' = By$. Note that there is a constant $Q_1 > 0$ so that 
\begin{equation*}
    \Vert x(t) \Vert = \Vert P(t)y(t) \Vert \leq \Vert P(t) \Vert \Vert y(t) \Vert  \leq Q_{1} \Vert y(t) \Vert \text{,}
\end{equation*}
for $t \in \mathbb{R}$ and since $y(t) = P^{-1}(t)x(t)$ there is a constant $Q_2 > 0$ such that 
\begin{equation*}
    \Vert y(t) \Vert = \Vert P^{-1}(t)x(t) \Vert \leq \Vert P^{-1}(t) \Vert \Vert x(t) \Vert  \leq Q_{2} \Vert x(t) \Vert \text{,}
\end{equation*}
for $t \in \mathbb{R}$. To conclude this proof, we must use some points from the Stability Theorem (2.49). The theorem states that \begin{enumerate}
    \item If $A$ has an eigenvalue with positive real part, then the trivial solution is unstable on $[0,\infty)$.
    \item If all the eigenvalues of $A$ with zero real parts are simple (multiplicity one) and all other eigenvalues of A have negative real parts, then the trivial solution is stable on $[0,\infty)$.
    \item If all the eigenvalues of $A$ have negative real parts, then the trivial solution of $x' = Ax$ is globally asymptotically stable on $[0,\infty)$.
\end{enumerate}

By applying the Stability Theorem to $B$ we can show that trivial solution's stability is as we claimed. In other words, we are mapping our original system to a linear system with constant real coefficients, $y' = B y$. Then, stability of our trivial solution is determined by the eigenvalues of $B$.

\section{Theorem 2.73}

In this theorem, we show some properties of Floquet multipliers as they relate to nontrivial solutions of a Floquet system. We also establish a condition for a nontrivial periodic solution to exist.

\subsection{Theorem}

Our claim is that $\mu_0$ is a Floquet multiplier of the Floquet system $x' = A(t)x$ if and only if there is a nontrivial solution $x$ such that $x(t + \omega) = \mu_0 x(t)$ for all $t \in \mathbb{R}$. Consequently, the Floquet system has a nontrivial periodic solution of period $\omega$ if and only if $\mu_0 = 1$ is a Floquet multiplier.

\subsection{Proof}

First, let us assume that $\mu_0$ is a Floquet multiplier of $x' = A(t)x$. Then, by definition $\mu_0$ must be an eigenvalue of 
\begin{equation*}
    C := \Phi^{-1}(0)\Phi(\omega)
\end{equation*}
where $\Phi$ is a fundamental matrix of $x' = A(t)x$. If we let $x_0$ be the eigenvector corresponding to $\mu_0$ and define a vector function $x(t) = \Phi(t)x_0$ with $t \in \mathbb{R}$, 
then we can see that $x$ is a nontrivial solution of $x' = A(t)x$ and \begin{equation*}
\begin{split}
    x(t + \omega) &= \Phi(t + \omega) x_0 \\
                  &= \Phi(t) C x_0\\
                  &= \Phi(t) \mu_0 x_0 \\
                  &= \mu_0 x(t) \\
\end{split}
\end{equation*}
for all $t \in \mathbb{R}$.

Conversely, we can also assume that there is a nontrivial solution $x$ such that \begin{equation*}
    x(t + \omega) = \mu_0 x(t)
\end{equation*}
for all $t \in \mathbb{R}$. Let $\Psi$ be a fundamental matrix of our Floquet system, then \begin{equation*}
    x(t) = \Psi(t) y_0
\end{equation*}
for all $t \in \mathbb{R}$ and some nontrivial vector $y_0$. Because Floquet's theorem says that the matrix function $\Psi(\cdot + \omega)$ is also a fundamental matrix, $x(t + \omega) = \mu_0 x(t)$, so $\Psi(t + \omega)y_0 = \mu_0 \Psi(t) y_0$. Therefore,\begin{equation*}
    \Psi(t) D y_0 = \Psi(t) \mu_0 y_0
\end{equation*}
where $D := \Psi^{-1}(0) \Psi(\omega)$. It follows that $D y_0 = \mu_0 y_0$, and so $\mu_0$ is an eigenvalue of $D$ and therefore a Floquet multiplier of our Floquet system.

\section{Closing Remarks}
Theorem 2.72 gave us qualitative information about the trivial solution. Specifically, we saw that the trivial solution is always globally asymptotically stable on $[0,\infty)$ and may be stable or unstable on $[0,\infty)$ depending upon the values of the Floquet multipliers. We accomplished this by linking Theorem 2.72 to the Stability Theorem, Theorem 2.49.

Theorem 2.73 showed us that $\mu_0$ is a Floquet multiplier of a Floquet system if and only if a nontrivial solution exists with $x(t+\omega) = \mu_0 x(t)$. We also showed that a nontrivial periodic solution exists if and only if $\mu_0 = 1$ is a Floquet multiplier. 

\end{document}

